---
title: "Reproducable Research Project 1"
author: "Ashton Chevallier"
date: "Thursday, April 16, 2015"
output: html_document
---

##Loading and preprocessing the data
1. Download and unzip data  
If you don't have in your wd, download it and unzip it.


```{r}
if(!file.exists("activity.csv")){
    if(!file.exists("activity.zip")){
        download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
                      ,destfile = "activity.zip")
    }
    unzip("activity.zip")
}

```
2. Read data in data frame using read.csv

```{r}

if(!exists("dataRaw")) {
    dataRaw <- read.csv("activity.csv")
}
```

3. Load packages and set NAs to blanks.

```{r, echo=FALSE,results='hide'}

library(ggplot2)
library(dplyr)
```
```{r}
#Transform Factor Dates to Date
days <- as.Date(dataRaw[,2])
steps <- dataRaw[,1]

#Create subset of NA Free data
data <- dataRaw[!is.na(dataRaw[,1]),]
```

###What is the total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
dailySteps <- group_by(data,date) %>% summarize(steps=sum(steps))
```
2. Make a histogram of total steps taken each day
```{r}
qplot(data$date, data=data, weight=steps, geom="histogram",binwidth = 1)
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(dailySteps$steps)
```

###What is the daily activity pattern
1. Time series plot of interval vs. avgsteps across all days in 
```{r}
avgSteps <- group_by(data,interval) %>% summarize(avgsteps=mean(steps))
qplot(data=avgSteps,x=interval,y=avgsteps,geom="line")
```

2. Find interval with maximum average steps
```{r}
intervalSteps <- group_by(data,interval) %>% summarize(totalsteps=sum(steps))
maxInterval <- which(intervalSteps$totalsteps==max(intervalSteps$totalsteps))
maxInterval
```

###Imputing Missing Values
1. Calculate the total missing values in the dataset (IE. total rows with NA)
```{r}
count(dataRaw[is.na(dataRaw),])
```
2. To fill in these missing values, we will be using the average daily steps calculated earlier  
   It's as good as method as any, and we already calculated it earlier.
3. Now we'll use a loop to fill in the data
```{r}
### Probably a for loop
```

4. Plot Histogram of new filled in data, then output the summary
```{r}
###Use GGPLOT again
```

###Are there differences in activity patterns between weekdays and weekends?
1. Create new variable weekday to the filled in dataset
```{r}
###Use Mutate and weekday()
```
2. Make panel plot of Weekdays vs. Weekends
```{r}
###Uge ggplot...need to look up the panel function though
```
