---
title: "Reproducable Research Project 1"
author: "Ashton Chevallier"
date: "Thursday, April 16, 2015"
output: html_document
---

##Loading and preprocessing the data
1. Download and unzip data  
If you don't have in your wd, download it and unzip it.


```{r}
if(!file.exists("activity.csv")){
    if(!file.exists("activity.zip")){
        download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
                      ,destfile = "activity.zip")
    }
    unzip("activity.zip")
}

```
2. Read data in data frame using read.csv

```{r}

if(!exists("dataRaw")) {
    dataRaw <- read.csv("activity.csv")
}
```

3. Load packages and set NAs to blanks.

```{r,results='hide'}

library(ggplot2)
library(dplyr)
```
```{r}
#Transform Factor Dates to Date
days <- as.Date(dataRaw[,2])
steps <- dataRaw[,1]

#Create subset of NA Free data
data <- dataRaw[!is.na(dataRaw[,1]),]
```

###What is the total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
dailySteps <- group_by(data,date) %>% summarize(steps=sum(steps))
```
2. Make a histogram of total steps taken each day
```{r}
qplot(data$date, data=data, weight=steps, geom="histogram",binwidth = 1)
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(dailySteps$steps)
```

###What is the daily activity pattern?
1. Time series plot of interval vs. avgsteps across all days in 
```{r}
avgSteps <- group_by(data,interval) %>% summarize(avgsteps=mean(steps))
qplot(data=avgSteps,x=interval,y=avgsteps,geom="line")
```

2. Find interval with maximum average steps
```{r}
intervalSteps <- group_by(data,interval) %>% summarize(totalsteps=sum(steps))
maxInterval <- which(intervalSteps$totalsteps==max(intervalSteps$totalsteps))
cbind(maxInterval,intervalSteps[maxInterval,])
```

###Imputing Missing Values
1. Calculate the total missing values in the dataset (IE. total rows with NA)
```{r}
count(dataRaw[is.na(dataRaw),])
```
2. To fill in these missing values, we will be using the average interval steps.
   It's as good as method as any, and we already calculated it earlier.
   
3. Now we'll use a loop to fill in the data.
```{r}
narows <- which(is.na(dataRaw))
zdata <-dataRaw
for(i in narows){
  zdata[i,1] <- avgSteps[which(avgSteps$interval==zdata[i,3]),2]
}
summary(is.na(zdata))
```

4. Plot Histogram of new filled in data, then output the summary
```{r}
qplot(zdata$date, data=zdata, weight=steps, geom="histogram",binwidth = 1)

summary(zdata$steps)
```

###Are there differences in activity patterns between weekdays and weekends?
1. Create new variable weekday to the filled in dataset
```{r,results ='hide'}
#Helper Function to make weekday or weekend
weekend <- function(x) {a
  if(x=='Sunday' | x=='Saturday') {y <- 'Weekend'}
  else {y <- 'Weekday'}
  y
  }

#Mutate to make new variable
zdata <- mutate(zdata, Weekday = weekdays(as.Date(date)))
zdata$Weekday <- sapply(zdata$Weekday,FUN=weekend)

```
2. Make panel plot of Weekdays vs. Weekends
```{r}
avgWeekdaySteps <- group_by(zdata,Weekday,interval) %>% summarize(avgsteps=mean(steps))
qplot(data=avgWeekdaySteps,x=interval,y=avgsteps,geom="line", facets = Weekday ~.)
```

